Male-biased sexual selection, but not sexual dichromatism, predicts speciation

### 2018-10-15 Sexual Selection and Speciation.Rmd
Just a few changes. check the numbering of the figs after s6.

We aimed to compile data on Passerine birds, which represent the largest group for which relatively complete plylogenetic, trait and environmental data are available. The following details the source and use of the data.

@@

To obtain speciation rate estimates we used Bayesian Analysis of Macroevolutionary Mixtures (BAMM). Additionally, we compared our BAMM output files (event data)  to an existing extinction rate dataset [@Harvey_2017].

@@

##Sexual Selection

Our proxy for sexual selection was based on measures of sexual dichromatism. We used two existing datasets; a complete set of male and female plumage scores in Passerines from @Dale_2015 and reflectance data from 1000 birds from @Armenta_2008.

the ratio of which provides an index of sexual dichromatism (SDi). The plumage score was scored using the _Handbook of the Birds of the World_ @delhoyo2011. By scanning images of males and females in this book across multiple patches, @Dale_2015 obtained mean plumage scores from RGB values. @Dale_2015 compared their RGB scores to a dataset of reflectance measurements of Australian birds. The two datasets were correlated but the relationship was expectedly noisy. Here we compare the @Dale_2015 data against another reflectance dataset from @Armenta_2008, which estimated sexual dichromatism as  the mean number of just noticeable differences (JNDs) between male and female plumage colour (discriminability), based on a model of bird colour vision.

Alongside sexual dichromatism, @Dale_2015 compiled available data for traits such as: Body size, tropical life history, Sexual selection, Cooperative breeding and Migration. We use a phylogenetic PCA (PC1) that estimates male-biased sexual selection from a species level dataset of sexual size dimorphism, social polygyny and paternal care.

@@

```{r, eval = FALSE}
#Obtain means
bird.frame <- lapply(bird.values, function(x) {as.data.frame(x)})
bird.summary <- lapply(bird.frame, function(x) {
(as.data.frame(apply(x,2,mean, na.rm =T )))})

@@

**4)** This process was repeated when we used gridded data for the LGM, LIG, NPP and human population density, while range size was obtained from the following code:

@@

##Spatial Variability PCAs

To obtain estimates of spatial variability we performed PCAs of the standard errors of the bioclimatic variables (excluding seasonality in temperature and precipitation). We expect variability to increase with increased range size. To correct for this association, we took the residuals from GAM models of the relationship between PCA components and range size.

**Table S1:** Loadings for the first two PCs, from a PCA on the variation (standard error, where n ~ 1,000) of each bioclim variable except the standard error of temperature seasonality (se.bioclim4) and the standard error of precipitation seasonality (se.bioclim15), as these are already measures of variability and based on other bioclim variables. We use the standard error as, while we attempted to sample each species range 1,000 times, some points returned NA values and would thus affect the standard deviation, but not standard error.
```{r}
restricted.data <- complete.dataframe %>% drop_na(bioclim1) #Drop species without environmental data
PCA.bioclim <- prcomp(restricted.data[c('se.bioclim1', 'se.bioclim2', 'se.bioclim3', 'se.bioclim5', 'se.bioclim6', 'se.bioclim7', 'se.bioclim8', 'se.bioclim9', 'se.bioclim10', 'se.bioclim11', 'se.bioclim12', 'se.bioclim13', 'se.bioclim14', 'se.bioclim16', 'se.bioclim17', 'se.bioclim18', 'se.bioclim19')], #Select se.bioclim
                      scale = TRUE, center = TRUE)
PCA.predictions <- predict(PCA.bioclim)
restricted.data <- cbind(restricted.data, PCA.predictions)

@@

```{r, fig.width=10, fig.height=5}
#run GAM
PC1.gam <- mgcv::gam(PC1*(-1) ~ s(log(range.size.m2)), data = restricted.data, family = "gaussian") #Inverse as the PC1 loads to the negative (COUNTER-NTUITIVE)

@@

**Figure S1:** The relationship between spatial variability in the temperature components (PC1) and log-range size is relatively strong. The relationship between spatial variability in precipitation (PC2) and log-range size is weaker. In the analysis we only used the residuals from PC1.

@@

To gain estimates for change in climate over the past ~130,000 years we used the difference in bioclim variables between the LIG and present values. The plots show the two PCs, broadly representing temperature and precipitation. Data is also provided for the last-glacial maximum (LGM) but was not used in the analysis.

@@

The sexual dimorphism dataset looks to have an over-dispersed distribution (see phylogenic plot in manuscript). Unfortunately, transformations do not greatly improve the distribution. Our model fit is expected to be reduced by this but further transformations are not obvious and would risk problems of interpretation.

@@

The functions that generate these tip rates estimates are: ``calculate.log.es``, ``calculate.nd`` and ``calculate.tb``. Here they are run on 100 trees plus a full MCC tree of passerine birds based on 2500 samples of the posterior [@Jetz_2012]. This was done using the ``phanghorn`` package [@phangorn_2010].

@@

```{r, fig.width=10, fig.height=6}
grid.newpage()
grid.draw(BAMM.variance)
```
**Figure S5:** The tip-rate estimate for BAMM are highly variable within each run of BAMM. Across most species there is high variability in the posterior distribution of tip-rate estimates. Here we show mean values, weights based on the variance and 50 % CIs.

@@

Given the high variability in tip-rate estimates from the above plot, below we peform some diagnostics on BAMM to demonstrate that the variability is unlikely an error in sampling or parameters, rather an inherent aspect of BAMM. The following is not inherently necessary to understand the conclusions drawn from the paper, however they do raise a set of methodological questions about BAMM that warrant further investigation.

@@

Notably, each of the top shifts has low probability, indicating that out of the entire posterior sample we cannot differentiate which shift configuration is more likely. Based on the following estimates of the BayesFactor, we are convinced that the number of shifts is non-zero, however it is still quite a wide distribution.

@@

```{r, fig.width = 8, fig.height = 5}
plotPrior(MCC.BAMM.mcmc, expectedNumberOfShifts=100)
```
**Figure S6:** The apparent convergence of the number of shifts in the posterior is at odds with the variability seen in the CSS and although there is greater certainty in the number of shifts being within the range above, the position of those shifts remains variable.

@@

# BAMM.variance <- ggExtra::ggMarginal(
#   p = BAMM.MCC.plot,
#   type = 'density',
#   margins = 'both',
#   size = 5,
#   colour = 'black',
#   fill = '#BA3B1C91'
# )
```
**Figure S7:** Based on the event data from the BAMM run by @Harvey_2017 we find that in both our case and theirs the tip-rate estimates for many species are extrememly variable. here we present mean estimates with 50 % CIs.

>>>>

We aimed to compile data on Passerine birds, which represent the largest group for which relatively complete phylogenetic, trait and environmental data are available. The following details the source and use of the data.

@@

To obtain speciation rate estimates we used Bayesian Analysis of Macroevolutionary Mixtures (BAMM). Additionally, we compared our BAMM output files (event data)  to an existing rate dataset [@Harvey_2017].

@@

##Sexual dichromatism and Male-biased Sexual Selection

The first proxy for sexual selection that we used was based on measures of sexual dichromatism. We used two existing datasets; a complete set of male and female plumage scores in Passerines from @Dale_2015 and reflectance data from 1000 birds from @Armenta_2008.

Briefly, @Dale_2015 obtained male and female plumage score, the ratio of which provides an index of sexual dichromatism (SDi). The plumage score was extracted using the _Handbook of the Birds of the World_ @delhoyo2011. By scanning images of males and females in this book across multiple patches, @Dale_2015 obtained mean plumage scores from RGB values. @Dale_2015 compared their RGB scores to a dataset of reflectance measurements of Australian birds. The two datasets were correlated but the relationship was expectedly noisy **include R square of correlation**. Here we compare the @Dale_2015 data against another reflectance dataset from @Armenta_2008, which estimated sexual dichromatism as  the mean number of just noticeable differences (JNDs) between male and female plumage colour (discriminability), based on a model of bird colour vision.

Alongside sexual dichromatism, @Dale_2015 compiled available data for traits such as: Body size, tropical life history, Sexual selection, Cooperative breeding and Migration. As a second measure of sexual selection, we used a phylogenetic PCA (PC1) that estimates male-biased sexual selection from a species level dataset of sexual size dimorphism, social polygyny and paternal care. Higher values indicate higher size dimorphism (larger males), higher polygyny and less paternal care.

@@

```{r, eval = FALSE}
#Obtain means per variable per species
bird.frame <- lapply(bird.values, function(x) {as.data.frame(x)})
bird.summary <- lapply(bird.frame, function(x) {
(as.data.frame(apply(x,2,mean, na.rm =T )))})

@@

**4)** The process above was repeated when we used gridded data for the LGM, LIG, NPP and human population density, while range size was obtained from the following code:

@@

##Environmental Spatial Variability PCAs

To obtain estimates of spatial variability in environment we performed PCAs of the standard errors of the bioclimatic variables across the 1000 random points extracted per species (excluding seasonality in temperature and precipitation). We expect variability to increase with increased range size. To correct for this association, and to be able to include both variables in the model, we took the residuals from GAM models of the relationship between PCA components and range size.

**Table S1:** Loadings for the first two PCs, from a PCA on the variation (standard error, where n ~ 1,000) of each bioclim variable except the standard error of temperature seasonality (se.bioclim4) and the standard error of precipitation seasonality (se.bioclim15), as these are already measures of variability and based on other bioclimatic variables. We use the standard error as, while we attempted to sample each species range 1,000 times, some points returned NA values and would thus affect the standard deviation, but not standard error.** mmm..SD is the sum of the distances of each point to the mean, divided by n-1, so it already takes into account whether some species have more or less points, it's like an average of distances to the mean. I think the standard way of measuring spread data is the SD more than the SE...anyway, I don't think that will change much if all Ns are very similar, but let's think about that for the publication**
```{r}
restricted.data <- complete.dataframe %>% drop_na(bioclim1) #Drop species without environmental data
PCA.bioclim <- prcomp(restricted.data[c('se.bioclim1', 'se.bioclim2', 'se.bioclim3', 'se.bioclim5', 'se.bioclim6', 'se.bioclim7', 'se.bioclim8', 'se.bioclim9', 'se.bioclim10', 'se.bioclim11', 'se.bioclim12', 'se.bioclim13', 'se.bioclim14', 'se.bioclim16', 'se.bioclim17', 'se.bioclim18', 'se.bioclim19')], #Select se.bioclim
                      scale = TRUE, center = TRUE)
PCA.predictions <- predict(PCA.bioclim)
restricted.data <- cbind(restricted.data, PCA.predictions)

@@

```{r, fig.width=10, fig.height=5}
#run GAM of association between range size and spatial climatic variation
PC1.gam <- mgcv::gam(PC1*(-1) ~ s(log(range.size.m2)), data = restricted.data, family = "gaussian") #Inverse as the PC1 loads to the negative (COUNTER-INTUITIVE)

@@

**Figure S1:** The relationship between spatial variability in the temperature components (PC1) and log-range size is relatively strong. The relationship between spatial variability in precipitation (PC2) and log-range size is weaker. In the analysis we only used the residuals from PC1. **How much % is explained by PC1??**

@@

Climate stability through time can potentially affect diversification dynamics. To gain estimates for change in climate over the past ~130,000 years we used the difference in bioclim variables between the LIG and present values. The plots show the two PCs, broadly representing temperature and precipitation. Data is also provided for the last-glacial maximum (LGM) but was not used in the analysis. **last sentence not needed, or otherwise explain why we chose LIG over LGM?**

@@

The sexual dimorphism dataset looks to have an over-dispersed distribution (see phylogenic plot in manuscript). Unfortunately, transformations do not greatly improve the distribution. Our model fit is expected to be reduced by this but further transformations are not obvious and would risk problems of interpretation. In any case, the phylogenetically corrected models we employ do not assume normality in the response variable.

@@

The functions that generate these tip rates estimates are: ``calculate.log.es``, ``calculate.nd`` and ``calculate.tb``. Here they are run on 100 trees plus a full MCC tree of passerine birds based on 2500 sample trees of the posterior [@Jetz_2012]. This was done using the ``phanghorn`` package [@phangorn_2010].

@@

```{r, fig.width=10, fig.height=6}
grid.newpage()
grid.draw(BAMM.variance)
```
**Figure S6:** **this changes to S6 and next to S7 etc**The tip-rate estimates for BAMM are highly variable within each run of BAMM. Across most species there is high variability in the posterior distribution of tip-rate estimates. Here we show mean values, weights based on the variance and 50 % CIs.

@@

Given the high variability in tip-rate estimates from the above plot, below we peformed some diagnostics on BAMM to demonstrate that the variability is unlikely an error in sampling or parameters, rather an inherent aspect of BAMM. The following is not inherently necessary to understand the conclusions drawn from the paper, however they do raise a set of methodological questions about BAMM that warrant further investigation.

@@

Notably, each of the top shifts has low probability, indicating that out of the entire posterior sample we cannot differentiate which shift configuration is more likely. Based on the following estimates of the BayesFactor, we are confident that the number of shifts is non-zero, however it is still quite a wide distribution.

@@

```{r, fig.width = 8, fig.height = 5}
plotPrior(MCC.BAMM.mcmc, expectedNumberOfShifts=100)
```
**Figure S7:** The apparent convergence of the number of shifts in the posterior is at odds with the variability seen in the CSS and although there is greater certainty in the number of shifts being within the range above, the position of those shifts remains variable.

@@

# BAMM.variance <- ggExtra::ggMarginal(
#   p = BAMM.MCC.plot,
#   type = 'density',
#   margins = 'both',
#   size = 5,
#   colour = 'black',
#   fill = '#BA3B1C91'
# )
```
**Figure S8:** Based on the event data from the BAMM run by @Harvey_2017 we find that in both our case and theirs the tip-rate estimates for many species are extremely variable across samples of the posterior probability. Here we present mean estimates with 50 % CIs.
